---
title: "Homework 6"
author: "Camille Okonkwo"
output: github_document
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
	echo = TRUE,
	fig.asp = 0.6,
	fig.width = 6,
	warning = FALSE,
	out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Set seed for reproducibility.

```{r seed}
set.seed(1)
```


###  Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modify victim_race to have categories (white and non-white), with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r homicide_df}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r baltimore_df}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r model_results_1}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New York. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race.

```{r model_results_2}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Problem 2

Downloading the Central Park weather data

```{r weather}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

We'll be looking at the distributions of these two quantities estimated from the weather_df data: r^2 and log(β̂ 1∗β̂ 2). 

Lets create a simple linear regression with `tmax` as the response and  `tmin` and `prcp` as predictors. 
```{r model}
model = lm(tmax ~ tmin + prcp, data = weather_df)
```


We're interested in the distribution of the quantities r^2 and log(β̂_1∗β̂_2) from weather_df. Let's extract r^2 and the beta_1 & beta_2 coefficients. 
```{r extract}
r_squared =
  model |> 
  broom::glance() |> 
  select(r.squared)

coefficients =
  model |> 
  broom::tidy() |>
  filter(term %in% c("tmin", "prcp")) |>
  select(term, estimate) |>
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  mutate(
    log_coefficient = ifelse(tmin > 0 & prcp > 0, log(tmin * prcp), NA)
    ) |> 
  select(log_coefficient)
```


It's important to point out that since β̂_prcp is negative, the log product of the coefficients will produce an Nan.

Using bootstrap sampling, we'll produce 5000 estimates of r^2 and log(β̂_1∗β̂_2). The most efficient way to do this would be utilizing a function, so let's go ahead and do that. 
```{r weather_boot}
weather_boot = function(df) {
  
  data = sample_frac(df, replace = TRUE)
  
r_squared =
    lm(tmax ~ tmin + prcp, data = data) |>
    broom::glance() |>
    select(r.squared)
  
  coefficients =
  model |> 
  broom::tidy() |>
  filter(term %in% c("tmin", "prcp")) |>
  select(term, estimate) |>
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  mutate(
    log_coefficient = ifelse(tmin > 0 & prcp > 0, log(tmin * prcp), NA)
    ) |> 
  select(log_coefficient)

  tibble(
    r_squared, coefficients
  )
}
# bootstrap 5000 times

weather_iter =
  tibble(strap_number = 1:5000) |> 
  mutate(
    strap_sample = map(strap_number, \(i) weather_boot(weather_df))) |> 
  unnest(strap_sample)
```

Now, we can plot the distribution r^2 and log(β̂_1∗β̂_2). 
```{r plots}

# distribution of r squared 
weather_iter |> 
  ggplot(aes(x = r.squared)) + 
  geom_density()

#distribution of log coefficients
weather_iter |> 
  ggplot(aes(x = log_coefficient)) +
  geom_density()
```

Out of 5000 values of log(β1*β2), ____ were removed from the analysis due to "N/A" values. 

The distribution of r^2 and log(β̂_1∗β̂_2) are approximately normal. 

With the 5000 bootstrap estimates, let's create a 95% Confidence Interval for r^2 and  log(β̂_1∗β̂_2). 
```{r weather_iter}
weather_iter |> 
  select(r.squared, log_coefficient) |> 
  pivot_longer(
    cols = c(r.squared, log_coefficient),
    names_to = "quantity", 
    values_to = "estimate"
  ) |> 
  group_by(quantity) |> 
  summarize(
    ci_LL = quantile(estimate, 0.025, na.rm = TRUE),
    ci_UL = quantile(estimate, 0.975, na.rm = TRUE)) |> 
  knitr::kable(digits = 4)
```

For r^2, the 95% Confidence Limit is (0.8885, 0.9407). The log coefficients confidence limit was not able to be computed due to the Nan values. 

### Problem 3

The goal of this problem is to understand the effects of several variables from birth-data on a child's birth weight. First, let's load and clean the data. 
```{r bw}
bw =
  read_csv("data/birthweight.csv") |> 
  janitor::clean_names() |> 
  mutate(
    babysex = as.factor(babysex), 
    frace = as.factor(frace), 
    malform = as.factor(malform), 
    mrace = as.factor(mrace)
  )
  
```

Let's create a linear regression model for this data. 

```{r model_birth}
model_birth =
  bw |> 
  lm(bwt ~ babysex + blength + gaweeks + fincome + wtgain, data = _)
```

I want to test if the sex of a child (`babysex`), the length of the child at birth (`blength`), the gestational age in weeks (`gaweeks`), family monthly income (`fincome`), and mother's weight gain (`wtgain`) are significant predictors of a child's birth weight. There is sound evidence that males are born at a higher weight than female children, and longer babies result in higher weighing babies. Older gestational age makes sense to contribute to higher weighing babies, as well as higher family incomes and higher amounts of weight gained by the mother carrying the child. 

Now, I will plot model residuals against predicted values. 
```{r residual_predicted}
bw |> 
  modelr::add_predictions(model_birth) |> 
  modelr::add_residuals(model_birth) |> 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() + 
  labs(title = "Residuals vs. Fitted Values",
       x = "Fitted Values",
       y = "Residuals") 
```

Let's compare the model I made to our two given models using `gather_predictions`.
```{r compare}
model_comp1= 
  bw |> 
  lm(bwt ~ blength + gaweeks, data = _)

model_comp2=
  bw |> 
  lm(bwt ~ babysex * bhead * blength, data = _)

bw |> 
  gather_predictions(model_birth, model_comp1, model_comp2) |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = blength, y = bwt)) + 
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red") + 
  facet_grid(~model)

```


Now, I'll make this comparison in terms of the cross-validation prediction error using `crossv_mc`. 
```{r cross_validation}
cross_val =
  crossv_mc(bw, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```


Let's fit the comparison models and assess prediction accuracy. We'll look at the RMSE's to do this.
```{r RMSE}
cross_val = 
  cross_val |> 
  mutate(
    model_mod  = map(train, \(df) lm(bwt ~ babysex + blength + gaweeks + fincome + wtgain, data = df)),
    comp1_mod     = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    comp2_mod  = map(train, \(df) lm(bwt ~ babysex*bhead*blength, data = df))) |> 
  mutate(
    rmse_model = map2_dbl(model_mod, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_comp1    = map2_dbl(comp1_mod, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_comp2 = map2_dbl(comp2_mod, test, \(mod, df) rmse(model = mod, data = df)))
```

Now, I'll plot the distribution of the RMSE models for each comparison model.
```{r RMSE_plot}
cross_val |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

The second comparison model with interactions has the lowest RSME distribution, so we can conclude that it has the best predictive accuracy. The RSME distribution for my proposed model and the first comparison model are similar, however my model has a slightly lower RMSE and lower minimum and maximum values. 
